Actuator line modeling of vertical-axis turbines

### 2016-05-12 paper.tex
Remove a bunch of DS model background info

DS models were first developed to improve predictive capability for helicopter
rotors, on which DS has significant effects on maneuverability and operational
envelope \cite{Bousman2000}. A summary of dynamic stall models developed for
helicopter rotors is presented in \cite{Leishman2006}. The simplest dynamic
stall models rely on semi-empirical correlations, e.g., the Gormont model
\cite{Gormont1973}, developed at the Boeing--Vertol Company. Several variants of
the Gormont model were developed for vertical-axis wind turbines, with varying
degrees of success; a summary is presented in \cite{Para2002}.

Leishman and Beddoes (LB) \cite{Leishman1989} developed a semi-empirical model
for unsteady aerodynamics and dynamic stall, which is derived from the
phenomenology of the physics instead rather than pure empiricism. Beddoes then
updated the model to the so-called third generation or ``3G'' version
\cite{Beddoes1993}. The LB DS models can be summarized conceptually based on the
following principles:
\begin{itemize}
    \item Dynamic conditions cause a time lag in effective angle of attack and
    lift force.

    \item Separation is determined by the Kirchoff flow approximation, which is
    also used to parameterize the normal force coefficient table based on the
    trailing edge separation point. This separation point also encounters a time
    lag.

    \item The separation initiates a vortex shedding cycle that causes an
    overshoot and subsequent undershoot in lift before returning to an attached
    flow condition.
\end{itemize}

Sheng et al. \cite{Sheng2008} developed an LB DS model variant targeted at low
Mach numbers. This model, along with the original and 3G LB DS model variants,
was tested for its effectiveness in cross-flow turbine conditions by Dyachuk et
al.~\cite{Dyachuk2014}, who concluded that the Sheng et al. variant results
matched most closely with experiments. In a similar study, the Sheng et al.
model also faired better than the Gormont model \cite{Dyachuk2015}, which
inspired its adoption here for the ALM.

>>>>

In this study we employed a dynamic stall model developed for low mach numbers
by Sheng et al. \cite{Sheng2008}---derived from the Leishman--Beddoes
semi-empirical model \cite{Leishman1989}. This model, along with two other
Leishman--Beddoes model variants, was tested for its effectiveness in cross-flow
turbine conditions by Dyachuk et al.~\cite{Dyachuk2014}, who concluded that the
Sheng et al. variant results matched most closely with experiments. In a similar
study \cite{Dyachuk2015}, the Sheng et al. model also performed better than the
Gormont model \cite{Gormont1973}, which inspired its adoption here for the ALM.
